### 2014-10-25 thesis/method/classification.tex
Finished section on feature selection and added references to equations

\section{0R and 1R}
Short for ``0-rule'' and ``1-rule'' respectively, 0R and 1R are
very simple classifiers. 0R uses no features to predict the class, and
1R only uses one.

@@

\section{C4.5 Decision Tree (DT)}
Decision trees (DT) can be specified recursively and are an example of a
``divide-and-conquer'' approach to learning. First, we choose a feature to
place at the root node and then create one branch for each possible value.
This partitions the data set into subsets, one for every value that the
feature takes. We repeat this for each branch, choosing different features
each time and continuing to divide the data set into progressively smaller
subsets, growing the tree.
If all examples in a subset have the same class at any point during
this process, we stop splitting that subset.

The only thing left to specify
now is how to decide which attribute to split on. Note that when splitting,
we would like to create subsets that are as ``pure''
-- that is, containing as many
instances of the same class -- as possible, so that the splitting may
terminate earlier. Since we would like our trees to be as small as possible to
prevent overfitting and allow the best generalisation ability,
we would therefore like to choose the feature to split on such that the
created subsets were as pure as possible. The measure of purity used in
learning decision trees is the \textit{information}, and is measured in
\textit{bits}. At any node (or subset) of the tree, this measure represents
the expected amount of information that would be needed to classify a new
example, given that the example reached the node. The information at each
node is calculated using the \textit{entropy} function:
\begin{equation*}
\mathrm{entropy}(p_1,p_2,\ldots,p_n) = -p_1\mathrm{log}p_1 -p_2\mathrm{log}p_2 \ldots -p_n\mathrm{log}p_n
\end{equation*}
Given $k$ classes, $c_1,\ldots,c_k$ is the number of instances in each
class at a node, and $\sum_{i=1}^k c_i = c$ is the total number of instances
at that node, the information of that node is then:
\begin{equation*}
\mathrm{info}(c_1,\ldots,c_k) = \mathrm{entropy}\left(\frac{c_1}{c},\frac{c_2}{c},\ldots,\frac{c_k}{c}\right)
\end{equation*}

@@

SVMs can be specified mathematically as a type of optimisation problem known
as \textit{constrained quadratic optimisation}. Given $m$ training vectors
$\mathbf{x_i} \in \mathbb{R}^n, i=1,\ldots,m$ and their labels
$y_i \in \{-1,+1\}, i=1,\ldots,m$, the optimisation problem can be written as:
\begin{equation}
\label{eqn:svm1}
\begin{aligned}
& \underset{\mathbf{w},b}{\mathrm{minimise}}
  && ||\mathbf{w}|| + C\sum_{i=1,\ldots,m} s_i \\
& \text{subject to}
  && y_i(\mathbf{w}^T\phi(\mathbf{x_i})+b) \geq 1 \\
& && s_i \geq 0, i = 1,\ldots,m
\end{aligned}
\end{equation}
where $\phi(\mathbf{x})$ is the kernel function, and
$\mathbf{w}\cdot\mathbf{x_i} + b = 0$ is the hyperplane that the SVM finds.
The $s_i$'s allow the SVM to ignore some examples if it cannot find a suitable
hyperplane, and $C$ (sometimes called the SVM's complexity constant) is a
constant specified by the user that controls how much we want to balance
finding the best possible margin with allowing misclassified examples.
Note that for a linear kernel, $\phi(\mathbf{x}) = \mathbf{x}$. Training the
SVM involves solving this optimisation problem and obtaining values for
$\mathbf{w}$ and $b$.

>>>>

\section{0R and 1R}
\label{sec:zeror}
Short for ``0-rule'' and ``1-rule'' respectively, 0R and 1R are
very simple classifiers. 0R uses no features to predict the class, and
1R only uses one.

@@

\section{C4.5 Decision Tree (DT)}
\label{sec:c45}
Decision trees (DT) can be specified recursively and are an example of a
``divide-and-conquer'' approach to learning. First, we choose a feature to
place at the root node and then create one branch for each possible value.
This partitions the data set into subsets, one for every value that the
feature takes. We repeat this for each branch, choosing different features
each time and continuing to divide the data set into progressively smaller
subsets, growing the tree.
If all examples in a subset have the same class at any point during
this process, we stop splitting that subset.

The only thing left to specify
now is how to decide which attribute to split on. Note that when splitting,
we would like to create subsets that are as ``pure''
-- that is, containing as many
instances of the same class -- as possible, so that the splitting may
terminate earlier. Since we would like our trees to be as small as possible to
prevent overfitting and allow the best generalisation ability,
we would therefore like to choose the feature to split on such that the
created subsets were as pure as possible. The measure of purity used in
learning decision trees is the \textit{information}, and is measured in
\textit{bits}. At any node (or subset) of the tree, this measure represents
the expected amount of information that would be needed to classify a new
example, given that the example reached the node. The information at each
node is calculated using the \textit{entropy} function:
\begin{equation}
\label{eqn:entropy}
\mathrm{entropy}(p_1,p_2,\ldots,p_n) = -p_1\mathrm{log}p_1 -p_2\mathrm{log}p_2 \ldots -p_n\mathrm{log}p_n
\end{equation}
Given $k$ classes, $c_1,\ldots,c_k$ is the number of instances in each
class at a node, and $\sum_{i=1}^k c_i = c$ is the total number of instances
at that node, the information of that node is then:
\begin{equation*}
\mathrm{info}(c_1,\ldots,c_k) = \mathrm{entropy}\left(\frac{c_1}{c},\frac{c_2}{c},\ldots,\frac{c_k}{c}\right)
\end{equation*}

@@

SVMs can be specified mathematically as a type of optimisation problem known
as \textit{constrained quadratic optimisation}. Given $m$ training vectors
$\mathbf{x_i} \in \mathbb{R}^n, i=1,\ldots,m$ and their labels
$y_i \in \{-1,+1\}, i=1,\ldots,m$, the optimisation problem can be written as:
\begin{equation*}
\begin{aligned}
& \underset{\mathbf{w},b}{\mathrm{minimise}}
  && ||\mathbf{w}|| + C\sum_{i=1,\ldots,m} s_i \\
& \text{subject to}
  && y_i(\mathbf{w}^T\phi(\mathbf{x_i})+b) \geq 1 \\
& && s_i \geq 0, i = 1,\ldots,m
\end{aligned}
\end{equation*}
where $\phi(\mathbf{x})$ is the kernel function, and
$\mathbf{w}\cdot\mathbf{x_i} + b = 0$ is the hyperplane that the SVM finds.
The $s_i$'s allow the SVM to ignore some examples if it cannot find a suitable
hyperplane, and $C$ (sometimes called the SVM's complexity constant) is a
constant specified by the user that controls how much we want to balance
finding the best possible margin with allowing misclassified examples.
Note that for a linear kernel, $\phi(\mathbf{x}) = \mathbf{x}$. Training the
SVM involves solving this optimisation problem and obtaining values for
$\mathbf{w}$ and $b$.

### 2014-10-25 thesis/method/selection.tex
Finished section on feature selection and added references to equations

Recall that each feature tells
us something about the training example. Ideally, it should also help us
classify the example. However, in practice and particularly
in medicine, a large number of attributes about a patient is usually recorded,
not all of which are relevant in predicting the class.
Having such a large number of features
is detrimental for two reasons: the training time is significantly
increased because the classifier must incorporate irrelevant information, and
this irrelevant information tends to degrade the performance
of even state-of-the-art classifiers such as C4.5 decision trees
\cite{Witten2005}. Clearly, there is an advantage in considering only a subset
of the most relevant features; to find them, we will use feature selection.

\section{Feature Selection}
After data preparation, selecting a subset of features or attributes to use
in a model is also important \citep{Kotsiantis2006},
and helps to reduce noise and redundant
attributes from becoming part of the final model. As we mentioned earlier,
experiments have found that
the performance of various classification methods such as decision trees
\textit{deteriorates} in the presence of useless attributes.
Our data set contains 73
non-target attributes, which is a fairly high number. Many of these attributes
are derived from others in the set: for example, whether or not a patient is
over 65 years of age (the \textit{age65} feature) is directly dependent on the
value of their \textit{age} attribute.

We began the process of feature selection by removing attributes that were
not available upon admission, as these tended to be highly related to the
LOS48 outcome and were a consequence, not a predictor, of LOS. These attributes
were: iculos, outcome, disp from ED, disp from hosp, died, rehab and los.
The id attribute was also removed because it simply numbered each record and
indicated nothing about the patient themself. This left us with 67 attributes.

The work of actually selecting the attributes out of the 67 available was
carried out in two ways: automated using feature selection methods, and
manually using advice from a domain expert.

\subsection{Automatic Methods}
There are a number of methods that are commonly used to automate the selection
of a subset of features for the final classifier. The methods we used were:
\begin{itemize}
  \item \textit{Correlation-based feature selection} \citep{Hall2000},
  which evaluates the `merit' of a
  subset of features by considering their degree of correlation with the target
  variable compared to their degree of correlation amongst each other. High
  correlation between variables indicates redundancy. Under
  this scheme, a set of features that are highly correlated with the target
  while having low correlation with other features in the set is ideal.
  \item Selection based on the
  \textit{Pearson product-moment correlation coefficient}, a measure of the
  linear correlation or dependence between two variables.\todo{add some more info}
  \item \textit{Information gain} \todo{explain a bit about this}
  \item Feature selection based on the \textit{OneR classifier}
  \citep{Holte1993}, which uses one
  attribute as the predictor for the target. This attribute is typically the one
  that has the lowest prediction error. Feature selection using this scheme
  involves first ranking attributes based on their error, with the lowest error
  attributes as the most desirable. We then selected the top 20 attributes to
  use in the building of the model.
  \item Wrapper feature evaluation \todo{explain}
\end{itemize}

We used the WEKA 3.7.11 \citep{Hall2009} implementation of the above
methods, after converting the data into the Attribute-Relation File Format
(ARFF)\footnote{\url{http://weka.wikispaces.com/ARFF+\%28developer+version\%29}}
most suitable for the program.
\todo{give more details, ie specifying binary/numeric
attributes, maybe specify in appendix?}

\subsection{Manual Selection}
We also wanted to compare the feature selection of an expert in the domain
of the data set, namely trauma, with the models obtained from the features
selected using the automated methods described above. This resulted in two
feature subsets. The first was obtained simply by presenting the entire list
of 73 features to the domain expert and asking him to select the features he
thought were most indicative of the target variable from experience. The
second set was obtained from the work of Dinh et. al. \citep{Dinh2013a} that
we are building on, giving us a baseline to compare our results to.

Altogether, the various feature selection methods resulted in a total of
7 data sets (4 automated, 2 expert selected, and 1 with all features) which
we used in the model construction phase.

>>>>

Recall that each feature tells
us something about a training example. Ideally, it should also help us
classify the example. However, in practice and particularly
in medicine, a large number of attributes about a patient is usually recorded,
not all of which are relevant in predicting the class.
Having such a large number of features
is detrimental for two reasons: the training time is significantly
increased because the classifier must incorporate more information than it
needs to, and
this irrelevant information tends to degrade the performance
of even state-of-the-art classifiers such as C4.5 decision trees
\cite{Witten2005}. Clearly, there is an advantage in considering only a subset
of the most relevant features; to find them, we will use what is called
\textit{feature selection}.

Our data set contains 78 features aside from the class, which is a fairly high
number. Many of these attributes
are derived from others in the set: for example, whether or not a patient is
over 65 years of age (the \texttt{age65} feature) is directly dependent on the
value of their \texttt{age} attribute.

\section{Preliminary Removal of Features}
We began the process of feature selection by removing attributes that were
not available upon admission, as these tended to be highly related to the
\texttt{los48} outcome and were a consequence, not a predictor, of LOS. Had
we not removed these features, they would be selected by the feature selection
methods described below due to their high association with \texttt{los48}, but
would actually tell us nothing about what causes a patient to require a
hospital stay of more or less than 2 days.
These attributes were: \texttt{iculos}, \texttt{outcome},
\texttt{disp from ed}, \texttt{disp from hosp}, \texttt{died}, \texttt{rehab},
\textit{los} and \texttt{eddc}.
The \texttt{id} attribute was also removed because it
simply numbered each record and
indicated nothing about the patient. This left us with 69 attributes.

We then removed 10 other features which were directly derived from others in
the data set: \texttt{gcs1}, \texttt{iss8}, \texttt{mechanism}, \texttt{male},
\texttt{age65}, \texttt{bp1}, \texttt{pr1}, \texttt{rr1}, \texttt{fall} and
\texttt{comorbidity}. At this stage there were 59 features -- still a large
number.

After eliminating redundant or inapplicable features as detailed above, we
considered four \textit{automatic} (correlation-based
feature selection, Pearson correlation, 1R and wrapper selection
using C4.5) and two \textit{manual} feature selection methods (domain expert
selection and the feature set used in existing work on trauma LOS prediction).
These methods are described in the following sections.

\section{Automatic Methods}
Automatic feature selection methods are useful for when we do not have a deep
understanding of the underlying domain and what each attribute really means to
the class. Even if we are well-versed in the domain, evaluating a few different
methods of automatic feature selection can give us new insight into the
learning problem at hand. Here we will describe a few of the methods we used.

\subsection{Correlation-based Feature Selection (CFS)}
Correlation-based feature selection (CFS) is a method developed by Hall
\cite{Hall2000} which looks for a subset of features that have high correlation
with the class but low intercorrelation among the features. In this case,
correlation between two features is measured using what is called the
\textit{symmetric uncertainty}:
\begin{equation*}
U(A,B) = 2\dfrac{H(A)+H(B)-H(A,B)}{H(A)+H(B)}
\end{equation*}
where $A$ and $B$ are two nominal attributes and $H$ is the entropy function
denoted by Equation \ref{eqn:entropy} in Chapter \ref{chap:classification}.
The entropies are calculated from the probabilities of each feature value, and
$H(A,B)$, the joint entropy of $A$ and $B$, is calculated from the
probabilities of each combination of values of $A$ and $B$. Note that $U(A,B)$
is always between 0 and 1.

To evaluate the ``goodness'' of a subset of features with the class $C$, we
compute:
\begin{equation}
\label{eqn:cfs}
\dfrac{\sum_j U(A_j,C)}{\sqrt{\sum_i \sum_j U(A_i,A_j)}}
\end{equation}
where $i$ and $j$ range over the features in the subset being evaluated. The
maximum value that Equation \ref{eqn:cfs} can obtain is 1 when all features
correlate perfectly with the class and also with each other -- the numerator
becomes $n$ and the denominator $\sqrt{n^2}$ if the subset has $n$ features.
\todo{elaborate a bit more}

\subsection{Pearson Correlation Coefficient}
The Pearson correlation coefficient measures the degree of linear correlation
between two variables. It is always in the range $[-1,1]$, where -1 is total
negative correlation, 0 is no (linear) correlation, and 1 is total positive
correlation. We compute the Pearson correlation between each feature and the
class using:
\begin{equation*}
r_i = \dfrac{\sum_{i=1}^m (x_i-\bar{x})(c-\bar{c})}{\sqrt{(\sum_{i=1}^m x_i-\bar{x})(\sum_{i=1}^m c-\bar{c})}}
\end{equation*}
where $m$ is the number of training examples, $x_i$ is the value of feature
$i$, and $\bar{x}$ and $\bar{c}$ are the arithmetic averages of the feature
values and the class values respectively. (Recall that we also used this
in Chapter \ref{chap:classification} to rank features for use in Ranked
Distance nearest neighbours.)

A higher value for $r_i$ indicates higher correlation with the class. We
rank all features according to their $r_i$ from highest to lowest, and select
some subset of these either by selecting a fixed number, or selecting all
features that have a correlation above a specified threshold.

\subsection{1R}
Another way to evaluate the relevance of features is to use the idea behind the
1R classifier described in Section \ref{sec:zeror}. Recall that the 1R
classifier constructs simple 1-rules based on each feature and uses the feature
that produces the lowest error rate (that is, the highest accuracy) on the data
set to classify unseen examples. These accuracy rates can be calculated for
each feature, which are then used to rank the features from highest to lowest
accuracy. The worst accuracy for any feature is the 0R accuracy, which is the
proportion of the majority class in the data set. This is because if 1R finds
that the accuracy of the rule constructed using a feature is worse than 0R,
then it will not use that rule and predict the majority class instead.

We can also select a fixed number of features or all features above a certain
1R accuracy threshold.

\subsection{Wrapper-based Feature Selection}
The feature selection methods described above (CFS, Pearson correlataion and
1R) have been described as \textit{filter} approaches to attribute selection.
Kohavi and John proposed another approach, the \textit{wrapper} approach:
instead of performing feature selection independently of the learning
algorithm using some measure of ``relevance'',
the learning algorithm itself is used to evaluate the goodness of
subsets of features \cite{Kohavi1997}. The difference between the filter
and wrapper approaches is depicted in Figure \ref{fig:wrapper}.

\begin{figure}[h]
\label{fig:wrapper}
\caption{}
\centering
%\includegraphics[scale=0.95]{images/a2-nn}
\end{figure}

A key advantage of the wrapper approach is that the selected feature sets are
specifically tailored to a learning algorithm, which ensures the best
performance of that algorithm given the data set. However, wrapper approaches
can also be prohibitively slow if the data or feature sets are large and the
chosen learning algorithm expensive to train (such as support vector machines
and multi-layer perceptrons). We will evaluate the use of C4.5 decision trees
(see Section \ref{sec:c45}) in selecting features using the wrapper approach.
Using decision trees to select features is relatively fast as due to the
divide and conquer nature of the learning algorithm, and the tendency of the
trees to select fewer features for learning than other algorithms.

\section{Manual Selection}
All of the above automatic feature selection methods use some notion of
``relevance'', such as symmetric uncertainty and linear correlation,
in order to determine the best subset of features to select.
Depending on the problem at hand, these definitions of relevance may or may
not work well with the data. In the LOS prediction literature, studies
have enlisted the help of medical domain experts to assist in feature
selection, due to their experience and understanding of the underlying
problem and the effect of the features on the outcome. It will therefore be
worthwhile to compare the performance of classifiers using features selected
automatically and manually.

\subsection{Baseline Feature Set}
The key previous work on trauma LOS prediction was by Dinh et. al.
\cite{Dinh2013a}, who used a specific feature set from the data set that we
use. We would like to use the features they selected as a baseline for
evaluating the other feature selection methods we describe.

\subsection{Domain Expert Selection}
Consultation with a domain expert yielded another feature set that we could
use to compare with the baseline and with the other automatic methods.

\section{Conclusion}
In this chapter we covered feature selection, which is concerned with reducing
the number of attributes that a learning algorithm needs to consider when
trying to learn a relationship between the training examples and their labels.
We described how we removed redundant and meaningless features before the
feature selection step, and discussed the automatic and manual feature
selection methods we use in our work. The value in using several methods of
each type of feature selection is two-fold: it allows us to not only compare
the performance of the methods, but also to compare the performance of the
different classifiers we use independently of the feature selection method.
